---
title: "Baltimore City Employees' Salaries"
output: 
html_notebook: default
---

1. Load the necessary package(s)
```{r}
library(tidyverse)
```

2. Load the data into a variable called "employees"
```{r}
employees <- read_csv("data/baltimore-city-employee-salaries.csv")
```
# Error: 'data/baltimore-city-employee-salaries.csv' does not exist in current working directory ('/Users/hannahjiang/Documents/github/hj-adj-summer22/assignments').

```{r setup}
knitr::opts_knit$set(root.dir = normalizePath("/Users/hannahjiang/Documents/github/hj-adj-summer22"))
```
```{r}
employees <- read_csv("data/baltimore-city-employee-salaries.csv")
```


## INTEGRITY CHECKS

Is any individual in here twice? Can you figure out why?
```{r}
employees %>% group_by(Name) %>% count() %>% arrange(desc(n));
employees %>% filter(Name == "Thomas,Stacey") %>% group_by(JobTitle) %>% count();
employees %>% filter(Name == "Carter,Angela") %>% group_by(JobTitle) %>% count();
employees %>% filter(Name == "Clark,Tony C") %>% group_by(JobTitle) %>% count();
employees %>% group_by(FiscalYear) %>% count()
```
# 1. They are different individuals with the same names; 2. the data was from 10 different fiscal years.

How many years of data do we have?
```{r}
employees %>% group_by(FiscalYear) %>% count()
```

What are the min and max annual salary given out in 2020? 
```{r}
employees %>% filter(FiscalYear == "FY2020") %>% arrange(desc(AnnualSalary));
employees %>% filter(FiscalYear == "FY2020") %>% arrange(AnnualSalary) %>% filter(AnnualSalary != 0) %>% arrange(AnnualSalary)
```

Which jobs get paid $0?
```{r}
employees %>% filter(AnnualSalary == 0) %>% group_by(JobTitle) %>% count()
```

How clean are the job titles? Are there a lot of duplicates? 
```{r}
employees %>% group_by(JobTitle) %>% count()
```

Clean up the JobTitles by making everything lowercase
(hint: use the mutate() function to overwrite the current JobTitle field, using the function str_to_lower())
```{r}
employees <- employees %>% mutate(JobTitle_Cleaned = str_to_lower(JobTitle))
```

Take a look at agency names; how clean are they? 
```{r}
employees %>% group_by(AgencyName) %>% count()
```

## QUESTIONS OF THE DATA

Who's the highest paid employee in FY2020?
```{r}
employees %>% filter(FiscalYear == "FY2020") %>% arrange(desc(GrossPay))
```


Which agency has the highest average salary in 2020? 
```{r}
employees %>%
  filter(FiscalYear == "FY2020") %>%
  group_by(AgencyName) %>%
  summarise(agency_avg_salary = mean(AnnualSalary)) %>% 
  arrange(desc(agency_avg_salary))
```

How many people work in the police department in 2020? 
```{r}
employees %>%
  filter(FiscalYear == "FY2020") %>% 
  filter(grepl("Police", AgencyName)) %>% 
  group_by(Name) %>% 
  count() %>% 
  arrange(desc(n));
employees %>% 
  filter(Name == "Pawley,Harry A")
```
# It seems that the two "Pawley,Harry A"s are different people. So the answer would be 3208 rows plus 1 = 3209.

How many are "police officers"?
```{r}
employees %>%
  filter(FiscalYear == "FY2020" & JobTitle_Cleaned == "police officer")
```

What is the police officers' total salary? Average salary?
```{r}
employees %>% 
  filter(FiscalYear == "FY2020" & JobTitle_Cleaned == "police officer") %>% 
  summarise(plof_tot_salary_2020 = sum(AnnualSalary), plof_avg_salary_2020 = mean(AnnualSalary))
```
