---
title: "Campaign Finance - Feb 28"
output: 
html_notebook: default
---
#Introduction: 
Each candidate for federal office (U.S. House of Representatives, U.S. Senate, President) must register with the Federal Election Commission (FEC) and form a Political Action Committee (PAC) to receive contributions. Candidates can have several different kinds of PACs, but everyone has a "principal campaign committee" through which they receive money.

It's an election year: every Representative is up for re-election in an even year, since Reps serve 2-year terms. One third of the Senate is up for re-election every two years because Sens serve 6-year terms. Candidates can receive money even if they're not running for re-election, but individuals are limited in how much they can give a candidate each 2-year cycle: $2900.

Missouri has 8 House seats and 2 Senate seats; only 1 Senate seat is up for re-election in 2022; Roy Blunt is not running for re-election so it's an open seat. Josh Hawley won't be up for re-election until 2024.

```{r}
library(tidyverse)
```

Load the three tables: contributions, candidates, committees
```{r}
contributions <- read_csv("data/campfin/mo_contributions.csv")
candidates <- read_csv("data/campfin/candidates.csv")
committees <- read_csv("data/campfin/committees.csv")
```

# Error, AGAIN: 'data/campfin/mo_contributions.csv' does not exist in current working directory ('/Users/hannahjiang/Documents/github/hj-adj-summer22/assignments').

# Some file names changed.

```{r setup}
knitr::opts_knit$set(root.dir = normalizePath("/Users/hannahjiang/Documents/github/hj-adj-summer22"))
```

```{r}
contributions <- read_csv("data/campfin/mo-contributions.csv")
candidates <- read_csv("data/campfin/candidates.csv")
committees <- read_csv("data/campfin/committees.csv")
```


#Documentation for these three tables: 

**Individual contributions**:
https://www.fec.gov/campaign-finance-data/contributions-individuals-file-description/
**Candidates**: 
https://www.fec.gov/campaign-finance-data/candidate-master-file-description/
**Committees**: 
https://www.fec.gov/campaign-finance-data/committee-master-file-description/


# JOINS
Join contributions and committees by cmte_id (the committee who filed the report, the receiving committee) 
```{r}
contributions %>% 
  inner_join(committees, by="cmte_id")
```

Join committees and candidates to see which candidate benefits from each committee (where applicable)
```{r}
committees %>% 
  inner_join(candidates, by="cand_id")
```


# QUESTIONS OF THE DATA

1. Which committee received the most money? Which candidate is associated with that committee? What office are they running for? 

```{r}
trsc_cmte <- inner_join(contributions, committees, by="cmte_id");
trsc_cand <- inner_join(trsc_cmte, candidates, by="cand_id");
trsc_cand %>% arrange(desc(transaction_amt))
```

2. How many candidates are running for the one open Senate seat in Missouri? 

```{r}
candidates %>% 
  filter(st == "MO" | city == ("ST LOUIS")) %>% 
  filter(office == "S") %>% 
  count()
```

3. Are Republican or Democratic candidates getting more money? 

```{r}
trsc_cand %>% 
  group_by(party_affil) %>% 
  summarise(party_amt = sum(transaction_amt)) %>% 
  arrange(desc(party_amt))
```

4. Are there any donors who work for the University of Missouri? 

```{r}
contributions %>% filter(grepl("UNIVERSITY OF MISSOURI", employer))
```

5. Which donor gave the most money?

```{r}
contributions %>%
  group_by(name) %>% 
  arrange(desc(transaction_amt))
```


# ASK AND ANSWER YOUR OWN THREE QUESTIONS

6. Which organization gave the most money, if a committee is connected to any organization?

```{r}
trsc_cmte %>% 
  group_by(connected_org_nm) %>% 
  summarise(org_amt = sum(transaction_amt)) %>% 
  arrange(desc(org_amt))
```

7. Which candidate was given the most money by Missourian donors?

```{r}
trsc_cand %>% 
  group_by(cand_name) %>% 
  filter(state == "MO") %>%
  summarise(mo_amt = sum(transaction_amt)) %>% 
  arrange(desc(mo_amt))
```

8. Which party was given the most money by homemakers?

```{r}
trsc_cmte %>% 
  group_by(party_affil) %>% 
  filter(grepl("HOMEMAKER", occupation) | grepl("WIFE", occupation) | grepl("SAHM", occupation) | grepl("MOTHER", occupation)) %>% 
  summarise(hmkr_cntr = sum(transaction_amt)) %>% 
  arrange(desc(hmkr_cntr))
```

